---
title: "Damage Function Analysis Report"
output:
  pdf_document: default
  html_notebook: default
---

# Data Summary

The dataset consists of regional polynomial coefficients transformed for analysis.

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(scales)
library(knitr)


# Set working directory
opts_knit$set(root.dir = "D:/CIL/damage_functions")
setwd("D:/CIL/damage_functions")

# Define input CSV file
outfile <- "results/collapse_FALSE_year_rcp_ssp_model_gcm_batch/analysis_scenarios/all_gamma_values/population_weighted/regional_polynomials.csv"


# Ensure the CSV file exists
if (!file.exists(outfile)) {
  stop("Error: The specified CSV file does not exist.")
}


# Load the data
results <- read.csv(outfile) %>%
  mutate(alpha = alpha * 1e5, beta = beta * 1e5, sigma11 = sigma11 * 1e10, 
         sigma12 = sigma12 * 1e10, sigma22 = sigma22 * 1e10, 
         zeta = zeta * 1e5, eta = eta * 1e5)

results.mu <- results[results$gamma == median(results$gamma),]


# Compute polynomial fits
TT <- seq(0, 20, length.out = 100)
if (nrow(results.mu) > 0) {
  allyy <- sapply(1:nrow(results.mu), function(ii) results.mu$alpha[ii] * TT + results.mu$beta[ii] * TT^2)
  pdf_data <- data.frame(
    TT, 
    yy = rowMeans(allyy, na.rm = TRUE), 
    q25 = apply(allyy, 1, function(yy) quantile(yy, .25, na.rm = TRUE)), 
    q75 = apply(allyy, 1, function(yy) quantile(yy, .75, na.rm = TRUE))
  )
} else {
  pdf_data <- data.frame(TT = numeric(), yy = numeric(), q25 = numeric(), q75 = numeric())
}
```

```{r data-summary, echo=TRUE}
summary(results.mu)
```


#

# Summarize the polynomials

```{r polynomial-fit, echo=FALSE, fig.height=4, fig.width=6}
if (nrow(pdf_data) > 0) {
  ggplot(pdf_data, aes(TT, yy)) +
    geom_line() + geom_ribbon(aes(ymin = q25, ymax = q75), alpha = .5) +
    xlab("Climatic temperature change") + ylab("Regional damage function") +
    theme_bw()
}
```


#

# Distribution of Zero Crossings


- Over what range of temperatures and how frequently do the polynomials dip below 0?
        - Crossing 0 at $$T = -\frac{\alpha}{\beta}$$.
        - $$\beta = 0$$ (no crossing): 15.51 %
        - $$T > 20$$ (beyond graph): 8.57 %
        - $$T < 0$$ (never crossing): 18.78 %
        
```{r}
# Calculate the percentage of cases for each condition
results.mu$iso3 <- substring(results.mu$region, 1, 3)

# Avoid division by zero by assigning NA where beta == 0
results.mu$cross0 <- ifelse(results.mu$beta == 0, NA, -results.mu$alpha / results.mu$beta)

# Compute proportions
prop_beta_zero <- mean(results.mu$beta == 0)  # Proportion where beta == 0
prop_cross0_above_20 <- mean(results.mu$alpha < 0 & results.mu$beta != 0 & results.mu$cross0 > 20, na.rm = TRUE)
prop_cross0_below_0 <- mean(results.mu$alpha > 0 & results.mu$beta != 0 & results.mu$cross0 < 0, na.rm = TRUE)

cat("Percentage where beta == 0 (No crossing):", round(100 * prop_beta_zero, 2), "%\n")
cat("Percentage where alpha < 0 and T > 20 (Beyond graph):", round(100 * prop_cross0_above_20, 2), "%\n")
cat("Percentage where alpha > 0 and T < 0 (Never crossing):", round(100 * prop_cross0_below_0, 2), "%\n")


```


```{r stats-crossings, echo=FALSE, fig.height=4, fig.width=6}
if (nrow(results.mu) > 0) {
  results.mu$iso3 <- substring(results.mu$region, 1, 3)
  results.mu$cross0 <- -results.mu$alpha / results.mu$beta
  ggplot(subset(results.mu, cross0 > 0 & cross0 < 20), aes(cross0)) +
    facet_wrap(~ ifelse(alpha > 0, "Concave", "Convex"), ncol = 1, scales = "free_y") +
    geom_histogram() +
    xlab("Temperature for crossing 0") +
    ylab("Region count") +
    theme_bw()
}
```


#

# Slope Analysis


How extreme are the slopes between 0 and 10 °C?
- Slope: $$\alpha + 2 \beta T$$, max at 0 or 10

```{r slope-analysis, echo=FALSE, fig.height=4, fig.width=6}
if (nrow(results.mu) > 0) {
  results.mu$maxslope <- pmax(results.mu$alpha + 2 * results.mu$beta * 10, results.mu$alpha, na.rm = TRUE)
  ggplot(results.mu, aes(maxslope)) +
    geom_histogram() + xlab("Maximum slope from 0 to 10°C") + ylab("Region count") +
    theme_bw()
}
```


When is concave polynomial estimated (and then clipped to linear)?
$$\beta > 0$$ (convex response): 84.46 %


Where are these?

```{r concave-poly, echo=FALSE, fig.height=4, fig.width=6}
results.mu$iso3 <- substring(results.mu$region, 1, 3)
results.mu$cross0 <- -results.mu$alpha / results.mu$beta

tbl <- table(results.mu$iso3[results.mu$beta == 0])
round(sapply(1:length(tbl), function(ii) tbl[ii] / sum(results.mu$iso3 == names(tbl)[ii])), 2)
```
  
  
#

# Model Fit


```{r load-model-data, echo=FALSE,results='hide'}
# Load data
opts_knit$set(root.dir = "D:/CIL/damage_functions")
setwd("C:/Users/scada/git/cil/dev/")
source("R/R/data_processing.R")


required_columns <- c("region", "year", "batch", "gcm", "model", "rcp", "ssp", "adjusted_mortality", "anomaly", "total_population", "gdppc")
csv_file_path <- "data/mortality_regression_full_mc.csv"

df <- load_and_process_data(csv_file_path, required_columns, gdp_baseline_start = 2010, gdp_baseline_end = 2020,
                            collapse_regional_data = FALSE)
```



```{r model-fit, echo=FALSE, fig.height=4, fig.width=6}
# Load necessary libraries
library(MASS) 
library(mvtnorm)

# Compute the median value of gamma from results
gamma.mu <- median(results$gamma)

# Initialize new columns in the dataframe
df$mean.normed <- df$delta_mortality / (exp(df$lgdp_delta)^gamma.mu)
df$mean.pred <- NA
df$mean.pred.mu <- NA
df$mean.pred.sd <- NA

# Get the total number of unique regions in the dataset
total_regions <- length(unique(df$region))

# Initialize a progress bar
pb <- txtProgressBar(min = 0, max = total_regions, style = 3)

# Get the list of unique regions
region_list <- unique(df$region)

# Loop over each region
for (i in seq_along(region_list)) {
  reg <- region_list[i]
  
  # Print progress message
  cat(sprintf("\nProcessing region: %s (%d/%d)\n", reg, i, total_regions))
  
  # Subset data for the current region
  subdf <- subset(df, region == reg)
  subres <- subset(results, region == reg & gamma == gamma.mu)
  
  # Check if either subdf or subres is empty, and skip if so
  if (nrow(subdf) == 0 || nrow(subres) == 0) {
    cat(sprintf("Skipping region: %s due to empty data\n", reg))
    next
  }
  
  # Compute the mean normalized values based on regression coefficients
  df$mean.normed[df$region == reg] <- subres$alpha * subdf$delta_temp + subres$beta * subdf$delta_temp^2
  df$mean.pred[df$region == reg] <- (subres$alpha * subdf$delta_temp + subres$beta * subdf$delta_temp^2) * exp(subdf$lgdp_delta)^gamma.mu

  # Subset results again for the region
  subres <- subset(results, region == reg)
  
  # Initialize a matrix to store 100 draws per row in subres
  draws <- matrix(NA, 100 * nrow(subres), nrow(subdf))
    
  # Loop over each row in subres to generate Monte Carlo simulations
  for (ii in 1:nrow(subres)) {
    # Generate random samples for alpha and beta using a multivariate normal distribution
    alphabetas <- rmvnorm(100, 
                           c(subres$alpha[ii], subres$beta[ii]), 
                           matrix(c(subres$sigma11[ii], subres$sigma12[ii], 
                                    subres$sigma12[ii], subres$sigma22[ii]), 2, 2))
    
    # Generate random error terms for the model
    thetas <- rnorm(100, 0, subres$zeta[ii])  # Error term related to temperature
    phis <- rnorm(100, 0, subres$eta[ii])  # General random error term
    
    # Compute estimates for each observation in the region
    ests <- sapply(1:nrow(subdf), function(jj) 
      (alphabetas[, 1] * subdf$delta_temp[jj] + alphabetas[, 2] * subdf$delta_temp[jj]^2) * 
      exp(subdf$lgdp_delta[jj])^subres$gamma[ii] + 
      thetas * subdf$delta_temp[jj] * exp(subdf$lgdp_delta[jj])^subres$gamma[ii] + 
      phis
    )
    
    # Store the simulated values
    draws[(1:100) + (ii - 1) * 100, ] <- ests
  }

  # Compute the mean and standard deviation of the predictions
  df$mean.pred.mu[df$region == reg] <- colMeans(draws, na.rm = TRUE)
  df$mean.pred.sd[df$region == reg] <- apply(draws, 2, sd, na.rm = TRUE)
  
  # Update the progress bar
  setTxtProgressBar(pb, i)
}

# Close the progress bar after completion
close(pb)



```
## vectorizing
```{r}
library(MASS)
library(mvtnorm)
library(parallel)  # for parallelization

# Compute the median value of gamma from results
gamma.mu <- median(results$gamma)

# Precompute common quantities outside the loop if possible
df$exp_lgdp_gamma <- exp(df$lgdp_delta)^gamma.mu

# Initialize new columns in the dataframe
df$mean.normed <- df$delta_mortality / (exp(df$lgdp_delta)^gamma.mu)
df$mean.pred <- NA
df$mean.pred.mu <- NA
df$mean.pred.sd <- NA

# Get the list of unique regions
region_list <- unique(df$region)
total_regions <- length(region_list)

# Define a function to process each region
process_region <- function(reg) {
  # Subset data for the current region
  subdf <- subset(df, region == reg)
  subres <- subset(results, region == reg & gamma == gamma.mu)
  
  if (nrow(subdf) == 0 || nrow(subres) == 0) {
    cat(sprintf("Skipping region: %s due to empty data\n", reg))
    return(NULL)
  }
  
  # Compute the mean normalized values based on regression coefficients
  # (Assuming subres has one row when filtering by gamma.mu for these predictions)
  df_mean_normed <- subres$alpha * subdf$delta_temp + subres$beta * subdf$delta_temp^2
  df_mean_pred <- df_mean_normed * subdf$exp_lgdp_gamma
  
  # Prepare to store Monte Carlo draws: one block per row in subres, each with 100 draws
  n_sims <- 100
  n_obs <- nrow(subdf)
  total_draws <- n_sims * nrow(subres)
  draws <- matrix(NA, nrow = total_draws, ncol = n_obs)
  
  # Create the design matrix for temperature effects
  X <- cbind(subdf$delta_temp, subdf$delta_temp^2)  # dimensions: n_obs x 2
  # The effect of GDP can be incorporated later by multiplying with subdf$exp_lgdp_gamma
  
  # For each row in subres, generate simulations in a vectorized way:
  for (ii in 1:nrow(subres)) {
    # Generate 100 draws for alpha and beta:
    cov_mat <- matrix(c(subres$sigma11[ii], subres$sigma12[ii],
                        subres$sigma12[ii], subres$sigma22[ii]), nrow = 2)
    draws_ab <- rmvnorm(n_sims, mean = c(subres$alpha[ii], subres$beta[ii]), sigma = cov_mat)
    # Compute the deterministic part: (alpha * temp + beta * temp^2)
    term1 <- draws_ab %*% t(X)  # dimensions: 100 x n_obs
    
    # Simulate error terms (vectorized over the 100 draws)
    theta <- rnorm(n_sims, mean = 0, sd = subres$zeta[ii])
    phi <- rnorm(n_sims, mean = 0, sd = subres$eta[ii])
    # The temperature error term: theta * delta_temp
    term2 <- outer(theta, subdf$delta_temp, "*")  # 100 x n_obs
    
    # Add the error terms; note that phi is added to every observation for a given draw
    sim_est <- (term1 + term2) * matrix(subdf$exp_lgdp_gamma, nrow = n_sims, ncol = n_obs, byrow = TRUE) +
      matrix(phi, nrow = n_sims, ncol = n_obs, byrow = FALSE)
    
    # Store these 100 draws in the appropriate rows of draws
    idx_start <- (ii - 1) * n_sims + 1
    idx_end <- ii * n_sims
    draws[idx_start:idx_end, ] <- sim_est
  }
  
  # Compute summary statistics for this region
  pred_mu <- colMeans(draws, na.rm = TRUE)
  pred_sd <- apply(draws, 2, sd, na.rm = TRUE)
  
  list(region = reg,
       mean.normed = df_mean_normed,
       mean.pred = df_mean_pred,
       mean.pred.mu = pred_mu,
       mean.pred.sd = pred_sd,
       indices = which(df$region == reg))
}

# Option A: Run sequentially (or use lapply)
results_list <- lapply(region_list, process_region)

# Option B: Run in parallel (adjust 'mc.cores' as appropriate)
# results_list <- mclapply(region_list, process_region, mc.cores = detectCores())




```
```{r}
# Remove any NULL results (regions that were skipped)
results_list <- Filter(Negate(is.null), results_list)

# Build a region-level summary data frame
# Construir un data frame con los resultados de cada región
df_sim <- do.call(rbind, lapply(results_list, function(res) {
  if (!is.null(res)) {
    data.frame(
      region         = res$region,
      mean_normed    = res$mean.normed,   # vector con tantos elementos como observaciones en la región
      mean_pred      = res$mean.pred,
      mean_pred_mu   = res$mean.pred.mu,
      mean_pred_sd   = res$mean.pred.sd,
      index          = res$indices
    )
  }
}))


# Si el data frame original tiene rownames o un identificador de fila
df_sim$index <- as.integer(df_sim$index)  # asegurarse que sean enteros

# Extraer del data frame original solo las filas que se procesaron
df_report <- df[unique(df_sim$index), ]
df_report$index <- as.numeric(rownames(df_report))

# Unir por el índice:
library(dplyr)
df_plot <- left_join(df_report, df_sim, by = "index")

```



```{r}
library(ggplot2)
library(scales)  # para usar muted()

ggplot(df_plot, aes(x = mean_normed, y = mean_pred )) +
  geom_point(size = 0.1, alpha = 0.1) +
  geom_abline(slope = 1, colour = muted('red')) +
  theme_bw() +
  xlab("Reported value") +
  ylab("Predicted value")

ggplot(df_plot, aes(x = mean_normed, y = mean_pred_mu)) +
  geom_point(size = 0.1, alpha = 0.1) +
  geom_abline(slope = 1, colour = muted('red')) +
  theme_bw() +
  xlab("Reported value") +
  ylab("Predicted value (mean of MCs)")


```







```{r}
df_region <- subset(df, region == 'CAN')

# Primer gráfico: Reported vs. Predicted
ggplot(df_region, aes(mean.normed, mean.pred)) +
    geom_point(size=.1, alpha=.1) + 
    geom_abline(slope=1, colour=muted('red')) +
    theme_bw() + 
    xlab("Reported value") + 
    ylab("Predicted value")

# Segundo gráfico: Reported vs. Predicted (mean of MCs)
ggplot(df_region, aes(mean, mean.pred.mu)) +
    geom_point(size=.1, alpha=.1) + 
    geom_abline(slope=1, colour=muted('red')) +
    theme_bw() + 
    xlab("Reported value") + 
    ylab("Predicted value (mean of MCs)")

```


